### 2019-12-26 docs/thesis-project/article.tex
updt

%%%%%%%%%%%%%%%%
% BEGIN HEADER %
%%%%%%%%%%%%%%%%
\title {Hives: File Survivability in P2P using Probabilistic Swarm Guidance
\thanks{This work has the support of the project MYRG2016-00097-FST of the University of Macau; by the Portuguese Fundação para a Ciência e a Tecnologia (FCT) through Institute for Systems and Robotics (ISR), under Laboratory for Robotics and Engineering Systems (LARSyS) project UID/EEA/50009/2019.}
}
\titlerunning{Hives}
\author{Francisco Barros, Daniel Silvestre \and Carlos Silvestre}
\authorrunning{F. Barros et al.}
\institute{Instituto Superior Técnico - Taguspark\newline Av. Prof. Doutor Cavaco Silva, 2744-016 Porto Salvo, Portugal
\email{fbarros@isr.ist.utl.pt, dsilvestre@isr.ist.utl.pt, csilvestre@umac.mo}\newline
\url{www.tecnico.ulisboa.pt}
}
\maketitle
%%%%%%%%%%%%%%
% END HEADER %
%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%
% BEGIN ABSTRACT %
%%%%%%%%%%%%%%%%%%
\begin{abstract}
Peer-to-Peer (P2P) systems have emerged as a potential choice to build large scale distributed storage systems, at a fraction of the cost of the alternative cloud approaches, which have a better quality of service guarantees. We survey general-purpose P2P protocols, distributed file systems, cloud-assisted P2P protocols, and erasure-code algorithms. We then draft a solution for the problem of creating a bio-inspired P2P overlay to be used on a cloud-assisted P2P storage and backup system using probabilistic swarm guidance. We anticipate that our solution will have high network overhead, in order to provide self-healing properties to the files stored in the system.

@@

\subsection{Problem Description}
With the growth of the internet and the emergence of distributed systems, two large-scale computing paradigms have gained popularity due to their promise of virtually unlimited scalability. On the one hand, we have Peer-to-Peer (P2P) networking, which can be defined as a group of equally privileged peers, who contribute with a portion of their resources, to achieve common goals. These networks are popular among file sharing and streaming applications \cite{ssaroiu:msp2pfss}, due to their self-organized behavior, lack of centralization and, low-cost. On the other hand, Cloud platforms offer unmatched, on-demand, self-service, availability, and reliability, at a higher cost, and are currently trendy, with companies such as Amazon, Google, and Microsoft offering various ITaaS products to individuals and organizations alike. Cloud-based systems are centralized architectures, in which a large number of computers are clustered and managed by master entities, which possibly, become bottlenecks. Both paradigms can be used to deploy distributed file-storage or file-backup applications; despite the fact that P2P implementations are cheaper for both companies and their clients, these have a hard-time achieving 99.9\% availability seen Cloud-based implementations, furthermore, depending on the P2P application there may also exist the risk of permanent file loss, making them somewhat unappealing for clients who want to store or backup their files remotely.

@@

%%%%%%%%%%%%%%%%%%%%%%
% BEGIN RELATED WORK %
%%%%%%%%%%%%%%%%%%%%%%
\newpage\section{Distributed File Systems}\label{sec:distributed-file-systems}
Distributed file system (DFS) enables individuals and organizations to store, access, and secure data on possibly remote locations, precisely as they would their local machines. It is common practice to replicate data at the file-level, block-level, or both, depending on the requirements of each system. For example, block-level replication may allow parallel execution of applications, such as is the case of Hadoop, Google, and Spark MapReduce algorithms. Here we present two DFS, which follow entirely different approaches when it comes to architecture and protocols. We focus only on the architectural implementation, their advantages or disadvantages. We do not consider parallel computation because our primary objective, in this thesis proposal, is the continuous availability of data in a system, to clients. Before we dive into them, it is essential to understand some desired characteristics of a DFS. In general, distributed systems of any kind should be \textbf{\textit{dependable}}, meaning that they should be able to avoid service failures that are more frequent and more severe than is acceptable. To be dependable is to be ready to provide the correct service (\textbf{\textit{availability}}), in a continuous manner (\textbf{\textit{reliability}}), with absence of catastrophic consequences to clients (\textbf{\textit{safety}}), e.g., losing a file, with no improper system alterations (\textbf{\textit{integrity}}) and, finally, to be able to undergo modifications and repairs at any time (\textbf{\textit{maintainability}}). Features that contribute to the above characteristics are summarized below.

\textbf{\textit{Load-Balancing}} is the process of balancing communication, processing, and storage overhead among all nodes of the system. It helps to avoid bottlenecks, increasing operational performance, and helps to prevent catastrophic failures. Prevention of failures is also achieved through \textbf{\textit{fault tolerance}}, which enables a system to continue operating, possibly at a reduced level,  rather  than failing completely,  when some part of it fails. As the amount of data passing in the network, being processed or stored in the system grows, or the number of clients using it increases, the capacity of a system to maintain its service without, or with little degradation, is called \textbf{\textit{scalability}}, it depends on the implementation of the system, and as a rule of thumb, centralized solutions are less scalable.

@@

\newpage\section{Block-Level Data Replication}
\newpage\section{Probabilistic Swarm Guidance}
Markov chains describe a sequence of possible events in which the probability of each event depends only on the previous event states, often used to model real-world processes, such as queues of customers arriving at an airport and population growths. Markov processes are the basis for stochastic simulation methods known as Markov Chain Monte Carlo (MCMC), which allow one to perform random sampling over vast state spaces, i.e., computing expectations concerning complex, high dimensional probability distributions. The idea behind MCMC is creating a Markov Chain, which will, over time, tend to an equilibrium that is close to desired density distribution. In robotics and control areas Markov Chains and MCMC are used as the basis foundation for a process known as probabilistic swarm guidance, which gifts autonomous robots with the capability of generating their trajectories independently of each other, in decentralized fashion and without ahead-of-time position allocation, so that sets of robots, known as swarm, converges to the desired distribution shape. Swarm guidance has the advantage over other methods for this problem because robots do not necessarily need to communicate with each other to perform their tasks, even if they could gain from it; their overall distribution can be adapted dynamically to the environment, and any damage to their desired formation is eventually self-repaired. Demonstration of these properties can be found in recent research by B. Açikmeçe \textit{et al.} \cite{psg-mca, psg-caa, psg-onoff}, which inspired us to apply swarm guidance principles to P2P networks and file survivability. Boyd et al. \cite{fast_mixing_mc} argue that the most popular MCMC methods, Metropolis-Hastings and Maximum-Degree Chain, that allow the creation of transition matrices, that can then be followed independently by autonomous robots in order to practice swarm guidance behavior are not optimal with respect to mixing rates, i.e.,  these algorithms may not result in matrices that converge in small amount of steps towards the desired density distribution. They propose that the problem of obtaining such matrices can be formulated as a convex optimization problem and prove that, for most cases, the resulting matrices have higher mixing rates. It is reasonable to assume that, concerning our problem in P2P networks, this might open the possibility of imposing additional constraints to the creation of the transition matrix, e.g., bounding the probability of a peer sending a file to some other peer in his hive to increase bandwidth balancing.

>>>>

%%%%%%%%%%%%%%%%
% BEGIN HEADER %
%%%%%%%%%%%%%%%%
\title {Hives: File Survivability in P2P using Probabilistic Swarm Guidance
\thanks{This work has the support of the project MYRG2016-00097-FST of the University of Macau; by the Portuguese Fundação para a Ciência e a Tecnologia (FCT) through Institute for Systems and Robotics (ISR), under Laboratory for Robotics and Engineering Systems (LARSyS) project UID/EEA/50009/2019.}
}
\titlerunning{Hives}
\author{Francisco Barros, Daniel Silvestre \and Carlos Silvestre}
\authorrunning{F. Barros et al.}
\institute{Instituto Superior Técnico - Taguspark\newline Av. Prof. Doutor Cavaco Silva, 2744-016 Porto Salvo, Portugal
\email{fbarros@isr.ist.utl.pt, dsilvestre@isr.ist.utl.pt, csilvestre@umac.mo}\newline
\url{www.tecnico.ulisboa.pt}
}
\maketitle
%%%%%%%%%%%%%%
% END HEADER %
%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%
% BEGIN ABSTRACT %
%%%%%%%%%%%%%%%%%%
\begin{abstract}
Peer-to-Peer (P2P) systems have emerged as a potential choice to build large scale distributed storage systems, at a fraction of the cost of the alternative cloud approaches, which have better quality of service guarantees. We survey general-purpose P2P protocols, distributed file systems, cloud-assisted P2P protocols, and erasure-code algorithms. We then draft a solution for the problem of creating a bio-inspired P2P overlay to be used on a cloud-assisted P2P storage and backup system using probabilistic swarm guidance. We anticipate that our solution will have high network overhead, in order to provide self-healing properties to the files stored in the system.

@@

\subsection{Problem Description}
With the growth of the internet and the emergence of distributed systems, two large-scale computing paradigms have gained popularity due to their promise of virtually unlimited scalability. On the one hand, we have Peer-to-Peer (P2P) networking, which can be defined as a group of equally privileged peers, who contribute with a portion of their resources, to achieve common goals. These networks are popular among file sharing and streaming applications \cite{ssaroiu:msp2pfss}, due to their self-organized behavior, lack of centralization and, low-cost. On the other hand, Cloud platforms offer unmatched, on-demand, self-service, availability, and reliability, at a higher cost, and are currently trendy, with companies such as Amazon, Google, and Microsoft offering various ITaaS products to individuals and organizations alike. Cloud-based systems are centralized architectures, in which a large number of computers are clustered and managed by master entities, which possibly, become bottlenecks. Both paradigms can be used to deploy distributed file-storage or file-backup applications; despite the fact that P2P implementations are cheaper for both companies and their clients, these have a hard-time achieving 99.9\% availability seen in Cloud-based implementations, furthermore, depending on the P2P application there may also exist the risk of permanent file loss, making them somewhat unappealing for clients who want to store or backup their files remotely.

@@

%%%%%%%%%%%%%%%%%%%%%%
% BEGIN RELATED WORK %
%%%%%%%%%%%%%%%%%%%%%%
\newpage\section{Distributed File Systems}\label{sec:distributed-file-systems}
Distributed file system (DFS) enables individuals and organizations to store, access, and secure data on possibly remote locations, precisely as they would their local machines. It is common practice to replicate data at the file-level, block-level, or both, depending on the requirements of each system. For example, block-level replication may allow parallel execution of applications, such as is the case of Hadoop, Google, and Spark MapReduce algorithms. Here we present two DFS, which follow entirely different approaches when it comes to architecture and protocols. We focus only on the architectural implementation, their advantages or disadvantages. We do not consider parallel computation because our primary objective, in this thesis proposal, is the continuous availability of data in a system, to clients. Before we dive into them, it is essential to understand some desired characteristics of a DFS. In general, distributed systems of any kind should be \textbf{\textit{dependable}}, meaning that they should be able to avoid service failures that are more frequent and more severe than is acceptable. To be dependable is to be ready to provide the correct service (\textbf{\textit{availability}}), in a continuous manner (\textbf{\textit{reliability}}), with absence of catastrophic consequences to clients (\textbf{\textit{safety}}), e.g., losing a file, with no improper system alterations (\textbf{\textit{integrity}}) and, finally, to be able to undergo modifications and repairs at any time (\textbf{\textit{maintainability}}). Some of the features that contribute to the above characteristics are summarized below.

\textbf{\textit{Load-Balancing}} is the process of balancing communication, processing, and storage overhead among all nodes of the system. It helps to avoid bottlenecks, increasing operational performance, and helps to prevent catastrophic failures. Prevention of failures is also achieved through \textbf{\textit{fault tolerance}}, which enables a system to continue operating, possibly at a reduced level,  rather  than failing completely, when some part of it fails, to this end, some sort of \textbf{\textit{machine redundancy}} and \textbf{\textit{data redundancy}} may be important. As the amount of data passing in the network, being processed or stored in the system grows, or the number of clients using it increases, the capacity of a system to maintain its service without, or with little degradation, is called \textbf{\textit{scalability}}, it depends on the implementation of the system, and as a rule of thumb, centralized solutions are less scalable.

@@

\newpage\section{Data Redundancy and File Survivability}\label{sec:dataredundancy}
In distributed storage systems, not losing files is of the utmost importance. Particularly, since peer availability in P2P networks is ever-changing and not guaranteed, the availability and even durability of a file may be endangered. A \textbf{\textit{durable file}} is one that, once put into the system, is not lost due to failures, whereas availability means that a user can access it promptly. For clarity, an available file is durable, but the other way is not necessarily correct. Distributed systems as a whole, usually ensure durability by employing file-level or block-level \textbf{\textit{replication}}; alternatively, some systems adopt \textbf{\textit{erasure coding}} techniques (\textit{coding} for short) or a mixture of both approaches. Replication has the advantage of having a more straightforward and less resource-intensive recovery model, on the other hand, coding has a better \textit{expansion factor}, i.e., can provide the same or even increased durability at a lower storage overhead. Examples already mentioned in this thesis proposal using replication include GFS and Hadoop, but there are many others; examples of modern systems using coding redundancy include Microsoft Azure and Facebook Analytics Hadoop Cluster \cite{lrc}. Because our proposal focuses on the survivability of files in P2P networks, using swarm guidance, where resources may be limited, we care only about coding in the survey that follows, as it allows more files to be kept by the system. We will briefly mention some of the raised concerns and inherent disadvantages associated with this technique and some of the suggested approaches to mitigate them over the two decades.

Two parameters specify some of the most widely used coding algorithms, the number of data symbols \textit{k} to be encoded, and the number of coded symbols \textit{n} to be produced, both assumed to be in a finite field $GF(2^w)$ and these are referred as \textit{(n, k) erasure-codes}. The fundamental idea, is that a file of size $S$ is divided into $k$ equally sized fragments and encode them into $n$ encoded fragments, allowing the original file to recovered from any set of encoded-fragment with count $k$. Reed-Solomon codes are one such example. Systems that use these and other more sophisticated coding are evaluated under some of the following metrics, the \textbf{\textit{repair-bandwidth}}, i.e., the number of bits communicated in the network during normal and repair operations, the \textbf{\textit{disk I/O}} which is the number of bits read from disk during each repair, the \textbf{\textit{repair locality}} which is the number of peers that would be involved in the repair process; Since coding algorithms are slow processes CPU cycles, XOR operation counts and cache misses are metrics often researched as well \cite{fast_coding}, when the main interest is tied with load-balancing and cost savings. We take particular interest in the first metric because, in general, coding leads to bursts in bandwidth consumption. Whenever a peer fails, all files he held need to be quickly recovered by some peer who needs to gather enough fragments from possibly many different peers. Such burst in consumption may lead to delays or even failure of the repair process due to physical link capacity between peers, who are involved in the repair process, or even throughput constraints enforced by each peer.

\newpage\section{Probabilistic Swarm Guidance}\label{sec:psg}
Markov chains describe a sequence of possible events in which the probability of each event depends only on the previous event states, often used to model real-world processes, such as queues of customers arriving at an airport and population growths. Markov processes are the basis for stochastic simulation methods known as Markov Chain Monte Carlo (MCMC), which allow one to perform random sampling over vast state spaces, i.e., computing expectations concerning complex, high dimensional probability distributions. The idea behind MCMC is creating a Markov Chain, which will, over time, tend to an equilibrium that is close to desired density distribution. In robotics and control areas Markov Chains and MCMC are used as the basis foundation for a process known as probabilistic swarm guidance, which gifts autonomous robots with the capability of generating their trajectories independently of each other, in decentralized fashion and without ahead-of-time position allocation, so that sets of robots, known as swarm, converges to the desired distribution shape. Swarm guidance has the advantage over other methods for this problem because robots do not necessarily need to communicate with each other to perform their tasks, even if they could gain from it; their overall distribution can be adapted dynamically to the environment, and any damage to their desired formation is eventually self-repaired. Demonstration of these properties can be found in recent research by B. Açikmeçe \textit{et al.} \cite{psg-mca, psg-caa, psg-onoff}, which inspired us to apply swarm guidance principles to P2P networks and file survivability. Boyd et al. \cite{fast_mixing_mc} argue that the most popular MCMC methods, Metropolis-Hastings and Maximum-Degree Chain, that allow the creation of transition matrices, that can then be followed independently by autonomous robots in order to practice swarm guidance behavior are not optimal with respect to mixing rates, i.e.,  these algorithms may not result in matrices that converge in small amount of steps towards the desired density distribution. They propose that the problem of obtaining such matrices can be formulated as a convex optimization problem and prove that, for most cases, the resulting matrices have higher mixing rates. It is reasonable to assume that, concerning our problem in P2P networks, this might open the possibility of imposing additional constraints on the creation of the transition matrix, e.g., bounding the probability of a peer sending a file to some other peer in his hive to increase bandwidth balancing.
